---
title: "prophet_package"
author: "Jeremy"
date: "6/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Prophet uses GAM framework:
yhat = trendFun + seasonal_trends + holidays + err


## Prophet function
```{r}
library(prophet)
df <- read.csv('C:/Users/jmeyer/Work_Code/prophet/example_wp_log_peyton_manning.txt') # timestamps are labeled ds
m1 <- prophet(df)

future <- make_future_dataframe(m1, periods=365) # Contains future dates
```

## Predictions
## Good indication of overfit/underfit
```{r}
forecast <- predict(m1, future)
tail(forecast[c('ds', 'yhat', 'yhat_lower', 'yhat_upper')])
# tail(forecast) # Also has seasonal compoents
```

## Plotting 
```{r}
plot(m1, forecast, ds='')
dyplot.prophet(m1, forecast) # Interactive
```


## PLotting seasonal components/trends
```{r}
prophet_plot_components(m1, forecast)

```


## Logistic Growth
### Note: The caps/floors need to be user-given. Implicit 0 floor
```{r}
df <- read.csv('examples/example_wp_log_R.csv')
df$cap <- 9
m <- prophet(df, growth = 'logistic')

future <- make_future_dataframe(m, periods = 1826)
future$cap <- 9
fcst <- predict(m, future)
plot(m, fcst)

```


```{r}
df2 <- df
df2$y <- 10 - df$y
df2$cap <- future$cap <- 6
df2$floor <- future$floor <- 1.5
m <- prophet(df2, growth = 'logistic')
fcst <- predict(m, future)
plot(m, fcst)
```


## Trend Changepoints. Sometimes TS experience abrupt changes in tregectories. These are added automatically only for first 80% of data 
```{r}
plot(m, fcst) + add_changepoints_to_plot(m)
dat <- cbind.data.frame(1:25,t(m$params$delta))
colnames(dat) <- c('Time', 'mag')
ggplot(dat) + geom_bar(aes(x=Time, y=mag), stat='identity') # Magnitude of changepoints (can be negative)

```

## The tuning parameter changepoint_prior_scale can also be fine tuned
## High values = More flexible and so more uncertanity
## Changepoints are assumed to continue at same rate as the past, so coverage intervals are not the same for every iteration and may not be accurate
```{r}
m <- prophet(df2, changepoint.prior.scale = 0.0005, n.changepoints = 10)
plot(m, fcst) + add_changepoints_to_plot(m)
ggplot(cbind.data.frame('Time'=1:10,'mag'=t(m$params$delta))) + 
  geom_bar(aes(x=Time, y=mag), stat='identity')

# Locations can be manually specified, then they will be regularized
```

## Special Dataes / Holidays
## Holidays are treated as indicator variables, each with its own parameter
## Windows around a specific holiday are treated as their own separate holiday
```{r}
library(dplyr)
playoffs <- data_frame(
  holiday = 'playoff',
  ds = as.Date(c('2008-01-13', '2009-01-03', '2010-01-16',
                 '2010-01-24', '2010-02-07', '2011-01-08',
                 '2013-01-12', '2014-01-12', '2014-01-19',
                 '2014-02-02', '2015-01-11', '2016-01-17',
                 '2016-01-24', '2016-02-07')),
  lower_window = 0,
  upper_window = 1
)
superbowls <- data_frame(
  holiday = 'superbowl',
  ds = as.Date(c('2010-02-07', '2014-02-02', '2016-02-07')),
  lower_window = 0,
  upper_window = 1
)
holidays <- bind_rows(playoffs, superbowls)
holidays

m3 <- prophet(df, holidays = holidays, holidays.prior.scale = 0.05) # Default value is 10. If you want more regularization, lower the scale. 
forecast3 <- predict(m3, future)

```
## Windows are to include days before/after event. Ex. Christmas eve

# Holiday Effect
```{r}
forecast3 %>% 
  select(ds, playoff, superbowl) %>% 
  filter(abs(playoff + superbowl) > 0) %>%
  tail(10)
prophet_plot_components(m3, forecast3)

```
## We can see what days / years have largest impact on views. We can also see spikes from the special hoildays.

## There's a built-in funciton that adds the common holidays for you
```{r}
m3 <- prophet(holidays = holidays)
m3 <- add_country_holidays(m3, country_name = 'US')
m3 <- fit.prophet(m3, df)
forecast3 <- predict(m3, future)
m3$train.holiday.names
prophet_plot_components(m3, forecast3)
```


## Fourier Order Seasonalities.
## N forier terms = @N variables. Be careful with overfitting
```{r}
df4 <- read.csv("examples/example_wp_log_peyton_manning.csv")
m4 <- prophet(df4)
prophet:::plot_yearly(m4)

m4 <- prophet(df4, yearly.seasonality = 20)
prophet:::plot_yearly(m4)

```
#### Default is 10 terms. If we have a more irregular pattern, we can increase the number of terms as so


## Custom Regular Seasonalities
```{r}
m5 <- prophet(weekly.seasonality=FALSE)
m5 <- add_seasonality(m5, name='monthly', period=30.5, fourier.order=5) # prior.scale can also be used here to regularize
m5 <- fit.prophet(m5, df)
forecast <- predict(m5, future)
prophet_plot_components(m5, forecast)
```


## Custom Conditional Seasonalities
## We have a seasonal effect that is different for the on/off seasons
```{r}
is_nfl_season <- function(ds) {
  dates <- as.Date(ds)
  month <- as.numeric(format(dates, '%m'))
  return(month > 8 | month < 2)
}
df4$on_season <- is_nfl_season(df4$ds)
df4$off_season <- !is_nfl_season(df4$ds) # Create conditionals

m6 <- prophet(weekly.seasonality=FALSE)
m6 <- add_seasonality(m6, name='weekly_on_season', period=7, fourier.order=3, condition.name='on_season')
m6 <- add_seasonality(m6, name='weekly_off_season', period=7, fourier.order=3, condition.name='off_season')
m6 <- fit.prophet(m6, df4)

future$on_season <- is_nfl_season(future$ds)
future$off_season <- !is_nfl_season(future$ds)
forecast <- predict(m6, future)
prophet_plot_components(m6, forecast)

```

## Overview. (Knobs): scale (fix overfitting), order (fix underfitting)
1. Changepoints (Locations, number of)
2. Holidays (Dates, regularization scale)
3. Seasonal Compoents (# Terms, length, scale)
4. Adding regressors (regressor, standardized)


## Adding regressors (don't have to be boolean values)
## Are put in as a linear component of the model
```{r}
nfl_sunday <- function(ds) {
  dates <- as.Date(ds)
  month <- as.numeric(format(dates, '%m'))
  as.numeric((weekdays(dates) == "Sunday") & (month > 8 | month < 2))
}
df$nfl_sunday <- nfl_sunday(df$ds)

m <- prophet()
m <- add_regressor(m, 'nfl_sunday')
m <- fit.prophet(m, df)

future$nfl_sunday <- nfl_sunday(future$ds)

forecast <- predict(m, future)
prophet_plot_components(m, forecast)
```


## Multiplicative Seasonality
## Problem: Series is not stationary. (Constant mean/variance)
## Addative seasonality does not work! Seasonlity is not a constant additive factor; it grows with the trend. (Increasing variance)
## Effects are in terms of percentages, not raw numbers
```{r}
df7 <- read.csv('examples/example_air_passengers.csv')
m7 <- prophet(df7)
future7 <- make_future_dataframe(m7, 50, freq = 'm')
forecast7 <- predict(m7, future7)
plot(m7, forecast7)


```

## Corrected
```{r}
m7 <- prophet(df7, seasonality.mode = 'multiplicative')
forecast7 <- predict(m7, future7)
plot(m7, forecast7)
```

## Components
```{r}
prophet_plot_components(m7, forecast7)

```


## Can mix/match time series effects
```{r}
# m <- prophet(seasonality.mode = 'multiplicative')
# m <- add_seasonality(m, 'quarterly', period = 91.25, fourier.order = 8, mode = 'additive')
# m <- add_regressor(m, 'regressor', mode = 'additive')
```


Width of uncertanity intervals by default is 80% (interval.width arguement)
Sources of Uncertanity
1. Trend
- Assumes that changepoints wil assume at same magnitude/frequency as the past (changepoint_prior_scale)
2. Seasonalities
- By default, these are turned off. Set mcmc.samples = N in prophet() to get components
3. Observation Noise

## Example
```{r}
m8 <- prophet(df, mcmc.samples = 300)
forecast8 <- predict(m8, future)
prophet_plot_components(m8, forecast8)
```

## Outliers
## Problem: Prophet will fit outliers using change points and it will expect the same occurances in the future -> LARGE Uncertanity intervals
```{r}
df9 <- read.csv('examples/example_wp_log_R_outliers1.csv')
m9 <- prophet(df9)
future9 <- make_future_dataframe(m9, periods = 1096)
forecast9 <- predict(m9, future9)
plot(m9, forecast9)
```


## Best way to deal with outliers is to remove them
## Prophet will replace NA with predicted values
```{r}
outliers <- (as.Date(df9$ds) > as.Date('2010-01-01')
             & as.Date(df9$ds) < as.Date('2011-01-01'))
df9$y[outliers] = NA
m <- prophet(df9)
forecast <- predict(m9, future9)
plot(m9, forecast9)
```

## Sometimes outliers can even mess up the predicted line
```{r}
df <- read.csv('examples/example_wp_log_R_outliers2.csv')
m <- prophet(df)
future <- make_future_dataframe(m, periods = 1096)
forecast <- predict(m, future)
plot(m, forecast)

outliers <- (as.Date(df$ds) > as.Date('2015-06-01')
             & as.Date(df$ds) < as.Date('2015-06-30'))
df$y[outliers] = NA
m <- prophet(df)
forecast <- predict(m, future)
plot(m, forecast)
```


## Diagnostics (Cross Validation)
# Horizion: Length of predicted window
# Period: Distance between each set cutoff values/bins
# Initial: How much of the training set to use before the cutoff value
# Default allows you to include sections (overlapping windows each containing 10% of data) of 
#     predictions to diagnose separately. 
```{r}
df.cv <- cross_validation(m, units='days', horizon=365, period=182, initial=1096)
performance_metrics(df.cv, rolling_window = 1) # Aggregated across all predictions
performance_metrics(df.cv, rolling_window = 0) # Error for each day after cutoff value
plot_cross_validation_metric(df.cv, metric = 'mae') # Blue line is a moving average taken over rolling window (10%)
plot_cross_validation_metric(df.cv, metric = 'mae', rolling_window = .1)

```
### Note: Make sure initial period is long enough to capture components. 
