---
title: "GGplot2"
author: "Jeremy"
date: "6/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Basic Scatterplot
```{r}
library(ggplot2)
g <- ggplot(midwest, aes(x=area, y=poptotal)) + 
       geom_point() + scale_y_continuous(labels = scales::comma)
g

```
## Aes() specifies axis values, but geom_point() actually graphs it
## the comma labels make the axis values more readable

# Draw line of best fit / Axis limits
```{r}
g + geom_smooth(method='lm', se=TRUE) + xlim(c(0, 0.1)) + ylim(c(0, 1000000))
g1 <- g + geom_smooth(method='lm', se=TRUE) + coord_cartesian(xlim=c(0,0.1), ylim=c(0, 1000000))  # zooms in without deleting points

plot(g1)
```
## Ggplot removed points outside ylimits before calculating line of best fit

## Title, subtitles and axis labels
```{r}
g1 + labs(title="Area Vs Population", subtitle="From midwest dataset", y="Population", x="Area", caption="Midwest Demographics")

# or

g1 + ggtitle("Area Vs Population", subtitle="From midwest dataset") + xlab("Area") + ylab("Population")
```

## Adjusting size /color of points
```{r}
ggplot(midwest, aes(x=area, y=poptotal)) + 
  geom_point(col="steelblue", size=3) +   # Set static color and size for points
  geom_smooth(method="lm", col="firebrick") +  # change the color of line
  coord_cartesian(xlim=c(0, 0.1), ylim=c(0, 1000000)) + 
  labs(title="Area Vs Population", subtitle="From midwest dataset", y="Population", x="Area", caption="Midwest Demographics")
```

## Adjusting Color based on categories
```{r}
gg <- ggplot(midwest, aes(x=area, y=poptotal)) + 
  geom_point(aes(col=state), size=3) +  # Set color to vary based on state categories.
  geom_smooth(method="lm", col="firebrick", size=2) + 
  coord_cartesian(xlim=c(0, 0.1), ylim=c(0, 1000000)) + 
  labs(title="Area Vs Population", subtitle="From midwest dataset",   y="Population", x="Area", caption="Midwest Demographics")
# + theme(legend.position="None")  # remove legend
plot(gg)
```
#### Wrap colors in aes() function. Describes how variables in data are mapped to visual properties. 

## 
```{r}
gg + scale_colour_brewer(palette = "Set1") # Can change entire clor set. 
# There are lots of palettes online  (RColorBrewer package)
```


```{r}
gg2 <- gg + scale_x_continuous(breaks=seq(0, 0.1, 0.01),labels = letters[1:11]) +
scale_y_continuous(breaks=seq(0, 1000000, 200000), labels = function(x){paste0(x/1000, 'K')})

gg2
```

## Setting Themes
```{r}
gg + theme_bw() + labs(subtitle="BW Theme")
gg + theme_classic() + labs(subtitle="Classic Theme")
```


Part 2:
```{r}
options(scipen=999) # No scientific notation
data("midwest", package = "ggplot2")

# Add plot components --------------------------------
gg <- ggplot(midwest, aes(x=area, y=poptotal)) + 
  geom_point(aes(col=state, size=popdensity)) + 
  geom_smooth(method="loess", se=F) + xlim(c(0, 0.1)) + ylim(c(0, 500000)) + 
  labs(title="Area Vs Population", y="Population", x="Area", caption="Source: midwest")

plot(gg)
```

##
```{r}
gg <- ggplot(midwest, aes(x=area, y=poptotal)) + 
  geom_point(aes(col=state, size=popdensity)) + 
  geom_smooth(method="loess", se=F) + xlim(c(0, 0.1)) + ylim(c(0, 500000)) + 
  labs(title="Area Vs Population", y="Population", x="Area", caption="Source: midwest")

gg + labs(color="State", size="Density")  # modify legend title

gg <- gg + guides(color=guide_legend("State"), size=guide_legend("Density"))  # modify legend title

gg + scale_color_discrete(name="State") + scale_size_continuous(name = "Density", guide = FALSE)  # turn off legend for size

```

## Manual Labeling of legend
```{r}
gg + scale_color_manual(name="State", 
                        labels = c("Illinois", 
                                   "Indiana", 
                                   "Michigan", 
                                   "Ohio", 
                                   "Wisconsin"), 
                        values = c("IL"="blue", 
                                   "IN"="red", 
                                   "MI"="green", 
                                   "OH"="brown", 
                                   "WI"="orange"))
```


## Legend Position
```{r}
# No legend --------------------------------------------------
gg + theme(legend.position="None") + labs(subtitle="No Legend")

# Legend to the left -----------------------------------------
gg + theme(legend.position="left") + labs(subtitle="Legend on the Left")

# legend at the bottom and horizontal ------------------------
gg + theme(legend.position="bottom", legend.box = "horizontal") + labs(subtitle="Legend at Bottom")

# legend at bottom-right, inside the plot --------------------
gg + theme(legend.title = element_text(size=12, color = "salmon", face="bold"),
           legend.justification=c(1,0), 
           legend.position=c(0.95, 0.05),  
           legend.background = element_blank(),
           legend.key = element_blank()) + 
  labs(subtitle="Legend: Bottom-Right Inside the Plot")

# legend at top-left, inside the plot -------------------------
gg + theme(legend.title = element_text(size=12, color = "salmon", face="bold"),
           legend.justification=c(0,1), 
           legend.position=c(0.05, 0.95),
           legend.background = element_blank(),
           legend.key = element_blank()) + 
  labs(subtitle="Legend: Top-Left Inside the Plot")
```

## Labels to a subset of data
```{r}
midwest_sub <- midwest[midwest$poptotal > 300000, ]
midwest_sub$large_county <- ifelse(midwest_sub$poptotal > 300000, midwest_sub$county, "")


gg + geom_text(aes(label=large_county), size=2, data=midwest_sub) + labs(subtitle="With ggplot2::geom_text") + theme(legend.position = "None")   # text

gg + geom_label(aes(label=large_county), size=2, data=midwest_sub, alpha=0.25) + labs(subtitle="With ggplot2::geom_label") + theme(legend.position = "None")  # label

# Plot text and label that REPELS eachother (using ggrepel pkg) ------------
library(ggrepel)
gg + geom_text_repel(aes(label=large_county), size=2, data=midwest_sub) + labs(subtitle="With ggrepel::geom_text_repel") + theme(legend.position = "None")   # text

gg + geom_label_repel(aes(label=large_county), size=2, data=midwest_sub) + labs(subtitle="With ggrepel::geom_label_repel") + theme(legend.position = "None")   # label
```


## Text to plot
```{r}
library(grid)
my_text <- "This text is at x=0.7 and y=0.8!"
my_grob = grid.text(my_text, x=0.7,  y=0.8, gp=gpar(col="firebrick", fontsize=14, fontface="bold"))
gg + annotation_custom(my_grob)
```

## Panels
```{r}
data(mpg, package="ggplot2")  # load data

g <- ggplot(mpg, aes(x=displ, y=hwy)) + 
      geom_point() + 
      labs(title="hwy vs displ", caption = "Source: mpg") +
      geom_smooth(method="lm", se=FALSE) + 
      theme_bw()  # apply bw theme
plot(g)

# Facet wrap with common scales
g + facet_wrap( ~ class, nrow=3) + labs(title="hwy vs displ", caption = "Source: mpg", subtitle="Ggplot2 - Faceting - Multiple plots in one figure")  # Shared scales

```


